---
title: "p8105_hw6_kdt2119"
author: Kelvin Tamakloe
date: December 4, 2021
output: github_document
---

```{r load_libraries, message=FALSE}
library(tidyverse)
library(purrr)
library(modelr)
library(viridis)
```

## Problem 1

**The code chunk below will perform the following function: Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).**

```{r import_bwt_data, message=FALSE}
birthweight_df = 
  read_csv("birthweight.csv") %>%
  mutate(babysex = factor(babysex, levels = c(1,2), labels = c("male", "female"))) %>%
  mutate(frace = factor(frace, levels = c(1,2,3,4,8,9), labels = c("White", "Black",
                        "Asian", "Puerto Rican", "Other", "Unknown"))) %>%
  mutate(malform = factor(malform, levels = c(0,1), labels = c("absent", "present"))) %>%
  mutate(mrace = factor(mrace, levels = c(1,2,3,4,8,9), labels = c("White", "Black",
                        "Asian", "Puerto Rican", "Other", "Unknown")))
colSums(is.na(birthweight_df))
```
The following variables were converted from numeric to factor variables: `babysex`, `frace`, `malform`, 
and `mrace`.

Further, the `birthweight_df` dataset has are no missing values (`NA` values).

**The code chunk below will perform the following function: Propose a regression model for birthweight.** 

The data was investigated to determine which variables may not be appropriate for inclusion in a proposed model.
 
```{r investigate_bwt, message=FALSE}
unique(pull(birthweight_df, pnumlbw))
unique(pull(birthweight_df, pnumsga))
```

* The `pnumlbw` (previous number of low birth weight babies) variable has all values = 0 and will as such not be an informative variable for inclusion in the model. 

* The `pnumsga` (number of prior small for gestational age babies) variable has all values = 0 and will as such not be an informative variable for inclusion in the model.

For my initial attempt, I included variables that,based on prior knowledge, I suspect are appropriate predictors of birthweight. These include: `babysex`, `bhead`, `blength`, `delwt`, `frace`,`gaweeks`, `malform`, `momage`, `mrace`, `ppbmi`, `smoken` and `wtgain`.


```{r model0, message=FALSE}
model0 = lm(bwt ~ babysex + bhead + blength + delwt + frace+ gaweeks + malform + momage + 
              mrace + ppbmi + smoken + wtgain, data = birthweight_df)
summary(model0)
```

The model's output suggests `frace`, `malform`, `momage`, and `wtgain` are not significant predictors of `bwt` (p-values > 0.05). These variables will be removed from the model and the model for `bwt` revised as below. 

```{r model1, message=FALSE}
model1 = lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + mrace + ppbmi + smoken, data = birthweight_df)
summary(model1)
model1 %>%
  broom::tidy()
```

This model's output shows that all of the variables included are significant predictors of `bwt`. The subcategory of `mrace` mother's race: Asian is not significant. However, the other subcategories of `mrace` are. As the variable is categorical, all levels would be included and remain in the model. 

**The code chunk below will perform the following function: Show a plot of model residuals against fitted values.**

```{r model_diagnostics, message=FALSE}
modelr::add_residuals(birthweight_df, model1)
modelr::add_predictions(birthweight_df, model1)
birthweight_df %>%
  modelr::add_residuals(model1) %>%
  modelr::add_predictions(model1) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point(color = "#21A8BD", alpha = 0.5) +
  labs(x = "Fitted Values", y = "Residuals") + 
  ggtitle("Plot of Model Residuals against Fitted Values") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 
```

**The code chunk below will perform the following function: Compare your model to two others in terms of the cross-validated prediction error**

* One using length at birth and gestational age as predictors (main effects only)

* One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r model_comparisons, message=FALSE}
cv_df =
  crossv_mc(birthweight_df, 50) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
cv_df2 = 
cv_df %>%
  mutate(
    model1 = map(train, ~lm(bwt ~ delwt + mrace + smoken + gaweeks + ppbmi + babysex + bhead + blength, data = .x)), 
    model2 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)), 
    model3 = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x))
  ) %>%
  mutate(
    rmse_model1 = map2_dbl(model1, test, ~rmse(model = .x, data = .y)), 
    rmse_model2 = map2_dbl(model2, test, ~rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(model3, test, ~rmse(model = .x, data = .y))
  )
cv_df2 %>% 
  select(
    starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
    ggplot(aes(x = model, y = rmse, fill = model)) +
    geom_boxplot() +
    scale_fill_viridis_d(option = "D") +
    labs(x = "Model", y = "RMSE", col = "Model") +
    ggtitle("RMSE Plots for All Models") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
```

From the plot, my model (model1) has a smaller RMSE than the RMSE of the other two models (model2 and model3). This shows that my model is a better fit for the data than the two other models. Model 2 (the model that without interaction terms) has the highest RMSE, suggesting that of the three models, it is the worst fitting model for the data. 